Near Optimal Behavior with Approximate Abstraction

### 2016-02-04 paper/example_domains.tex
example rewritten, results rewritten

\section{Example Domains}
To demonstrate the efficacy of these approximate abstraction functions, we test functions in the $\ep{Q^*}$ abstraction family on four domains -- NChain, Taxi, Minefield and Random MDP, which we dub ``Random". These domains were selected for the variety they offer -- Nchain is relatively simple domain, Taxi is goal-based and hierarchical in nature, Minefield is stochastic and Random MDP has a high number of near-optimal behavior.

@@

NChain, pictured in Figure~\ref{fig:nchain-visual}, is a simple MDP heavily investigated in the Bayesian Reinforcement Learning literature due to the interesting exploration problem it poses. In all states, the agent has two actions available: advance down the chain, or return to state 0. The agent receives a small positive reward, $r_{small}$ for returning to state 0, and no reward for advancing down the chain, unless the agent transitions to the last state in the chain. This transition provides a large positive reward, $r_{large}$. Thus, the agent must explore down the chain and avoid the intermediate smaller rewards in order to find the large reward at the end. In many implementations, there is also a small slip probability that the applied action has the unintended effect of the opposing action (so advancing down the chain with probability $\rho$ will take the agent back to state 0, and vice versa). In our implementation, we set $N=10$, $\rho=0.2$, $r_{small} = 0.2$, and $r_{large} = 1.0$.

@@

Taxi has long been studied by the hierarchical reinforcement learning literature due to its natural decomposition into hierarchical subgoals. The agent, operating in a grid world style domain, may move left, right, up, and down, as well as pickup a passenger and drop off a passenger. The agent must seek out passengers and take them to their destination. We visualize the compression on a simple 626 Taxi instance in Figure~\ref{fig:taxi-visual}. As stated above, we visualize the original Taxi problem into a graph representation so that we may visualize both the ground MDP and abstract MDP in the same format.

% Taxi Compression Visuals
\begin{figure}[h]
\subfigure[Ground Taxi]{
\includegraphics[width=0.48\columnwidth]{figures/ground_taxi.png}}
\subfigure[Abstract Taxi]{
\includegraphics[width=0.48\columnwidth]{figures/abs_taxi.png}}
\label{fig:taxi-visual}
\caption{Comparison of Original Taxi MDP and Abstract Taxi MDP, under an $\epQ$ abstraction, with $\epsilon=0.03$}
\end{figure}

@@

% Minefield Compression Visuals
\begin{figure}
\subfigure[Ground Minefield]{
\includegraphics[width=0.48\columnwidth]{figures/ground_minefield.png}}
\subfigure[Abstract Minefield]{
\includegraphics[width=0.48\columnwidth]{figures/abs_minefield.png}}
\label{fig:minefield-visual}
\caption{Comparison of Original Minefield MDP and Abstract MDP, under an $\epQ$ abstraction, with $\epsilon=.8$}
\end{figure}

@@

The Random MDP domain we consider is as follows. There are $N$ states, and $K$ actions. For each state, each action can transition to two randomly selected (but fixed) states, each with probability $0.5$. The Random MDP and its compression are visualized in Figure~\ref{fig:minefield-visual}.

% Random Compression Visuals
\begin{figure}
\subfigure[Ground Random]{
\includegraphics[width=0.48\columnwidth]{figures/ground_random.png}}
\subfigure[Abstract Random]{
\includegraphics[width=0.48\columnwidth]{figures/abs_random.png}}
\label{fig:minefield-visual}
\caption{Comparison of Original Random MDP and Abstract Random MDP, under an $\epQ$ abstraction, with $\epsilon=0.2$}
\end{figure}

>>>>

\section{Example Domains}
To demonstrate the efficacy of these approximate abstraction functions, we test functions in the $\ep{Q^*}$ abstraction family on four domains -- NChain, Taxi, Minefield and Random MDP, which we dub ``Random". These domains were selected for their variety -- NChain is relatively simple, Taxi is goal-based and hierarchical in nature, Minefield is stochastic, and Random MDP has many near-optimal policies.

@@

NChain, pictured in Figure~\ref{fig:nchain-visual}, is a simple MDP heavily investigated in the Bayesian Reinforcement Learning literature due to the interesting exploration problem it poses~\cite{dearden1998bayesian}. In all states, the agent has two actions available: advance down the chain, or return to state 0. The agent receives a small positive reward, $r_{small}$ for returning to state 0, and no reward for advancing down the chain, unless the agent transitions to the last state in the chain. This transition provides a large positive reward, $r_{large}$. Thus, the agent must explore down the chain and avoid the intermediate smaller rewards in order to find the large reward at the end. Transitions also have small slip probability such that the applied action has the unintended effect of the opposing action (so advancing down the chain with probability $\rho$ will take the agent back to state 0, and vice versa). In our implementation, we set $N=10$, $\rho=0.2$, $r_{small} = 0.2$, and $r_{large} = 1.0$.

@@

Taxi has long been studied by the hierarchical reinforcement learning literature due to its natural decomposition into hierarchical subgoals~\cite{dietterich2000hierarchical}. The agent, operating in a grid world style domain, may move left, right, up, and down, as well as pickup a passenger and drop off a passenger. The goal is achieved when the agent has taken each passenger to its destination.

%We visualize the compression on a simple 626 Taxi instance in Figure~\ref{fig:taxi-visual}. As stated above, we visualize the original Taxi problem into a graph representation so that we may visualize both the ground MDP and abstract MDP in the same format.

% Taxi Compression Visuals
%\begin{figure}[h]
%\subfigure[Ground Taxi]{
%\includegraphics[width=0.48\columnwidth]{figures/ground_taxi.png}}
%\subfigure[Abstract Taxi]{
%\includegraphics[width=0.48\columnwidth]{figures/abs_taxi.png}}
%\label{fig:taxi-visual}
%\caption{Comparison of Original Taxi MDP and Abstract Taxi MDP, under an $\epQ$ abstraction, with $\epsilon=0.03$}
%\end{figure}

@@

% Minefield Compression Visuals
\begin{figure}
\subfigure[Ground Minefield]{
\includegraphics[width=0.48\columnwidth]{figures/ground_minefield.png}}
\subfigure[Abstract Minefield]{
\includegraphics[width=0.48\columnwidth]{figures/abs_minefield.png}}
\label{fig:minefield-visual}
\caption{Comparison of Original Minefield MDP and Abstract MDP, under an $\epQ$ abstraction, with $\epsilon=1.1$}
\end{figure}

@@

The Random MDP domain we consider is as follows. There are $N$ states, and $K$ actions. For each state, each action can transition to two randomly selected (but fixed) states, each with probability $0.5$.

%The Random MDP and its compression are visualized in Figure~\ref{fig:minefield-visual}.

% Random Compression Visuals
%\begin{figure}
%\subfigure[Ground Random]{
%\includegraphics[width=0.48\columnwidth]{figures/ground_random.png}}
%\subfigure[Abstract Random]{
%\includegraphics[width=0.48\columnwidth]{figures/abs_random.png}}
%\label{fig:minefield-visual}
%\caption{Comparison of Original Random MDP and Abstract Random MDP, under an $\epQ$ abstraction, with $\epsilon=0.2$}
%\end{figure}

### 2016-02-04 paper/results_discussion.tex
Intro done for now
example rewritten, results rewritten

\section{Approximate Abstraction on Example Domains}
%Explanation of what abstraction is used
We run experiments on aggregation functions the $\epQ$ abstraction. For each trial our aggregation greedily groups ground states into abstract states that satisfy the $\epQ$ criteria. The order in which ground states are selected and then greedily added is randomized so there is randomization across trials. Every ground state is equally weighted in its abstract state.

%Explanation of what plots done
For each domain we report two quantities as a function of epsilon with \%95 confidence bars. First, we compare the number of states in the abstract \ac{MDP} to different values of $\epsilon$, shown in Figure~\ref{fig:main_empirical_results1}. The smaller this value, the smaller the state space of the $\ac{MDP}$ that the agent must plan over. Second, we report the value under the abstract policy of the initial ground state, shown in Figure~\ref{fig:main_empirical_results2}. 200 Trials are run for each data point.

%Discuss results
Our empirical results corroborate our thesis -- that approximate state abstractions can form the basis of learnable and useful state aggregation criteria. In both NChain and Minefield, we observe that despite increasing $\epsilon$ to reduce the number of states which must be planned over, optimal behavior is either fully maintained (NChain) or very nearly maintained (Minefield). Similarly for Taxi and $\epsilon$ between $.02$ and $.025$ we observe a reduction in the number of states in the Abstract MDP while value is fully maintained. After $.025$, increased reduction in state space size comes at a cost of value. Lastly, as $\epsilon$ is increased in the Random domain, there is a smooth reduction in the number of abstract states which must be planned over with a corresponding cost in the value of the derived policy.

Our experimental results also highlight a noteworthy characteristic of approximate state abstraction in goal-based \acp{MDP}. Taxi exhibits relative stability in state space size and behavior for epsilon up to $.02$ until both fall off dramatically. We attribute the sudden fall off in number of abstract states and value of the derived policy in Taxi to its goal-based nature; once information critical for achieving optimal behavior is lost in the state aggregation, solving the goal -- and so acquiring any reward -- is unattainable. Conversely, in the Random domain, a great deal of near-optimal policies are available to the agent. Thus, even the information for optimal behavior is lost with an increase in $\epsilon$ there are a great deal of near-optimal policies available to the agent which are not lost in the state aggregation.

\enote{Should we regenerate plots for nChain so it has CIs to match other plots?}
% Figure: Epsilon vs. #States and Epsilon vs Abstract Pol. For NChain and Minefield.
\begin{figure}
\label{fig:main_empirical_results1}
\centering

>>>>

\section{Approximate Abstraction on Example Domains}
%Explanation of what abstraction is used
We run experiments on aggregation functions in the $\epQ$ family of approximate abstractions. For each trial our aggregation greedily groups ground states into abstract states that satisfy the $\epQ$ criteria. The order in which ground states are added is randomized, so there is randomization across trials. Every ground state is equally weighted in its abstract state.

%Explanation of what plots done
For each domain we report two quantities as a function of epsilon with \%95 confidence bars. First, we compare the number of states in the abstract \ac{MDP} to different values of $\epsilon$, shown in Figure~\ref{fig:main_empirical_results1}. The smaller this value, the smaller the state space of the $\ac{MDP}$ that the agent must plan over. Second, we report the value under the abstract policy of the initial ground state, shown in Figure~\ref{fig:main_empirical_results2}. In the Taxi and Random domains, 200 trials are run for each data point, whereas 20 trials are run for data points in Minefield and NChain.

%Discuss results
\enote{Make sure results discussion is consistent with "the thesis of this work..."}

Our empirical results corroborate our thesis -- that approximate state abstractions can form the basis of learnable and useful state aggregation criteria. In both NChain and Minefield, we observe that as $\epsilon$ increases from $0$, the number of states which must be planned over is reduced, and optimal behavior is either fully maintained (NChain) or very nearly maintained (Minefield). Similarly for Taxi, when $\epsilon$ is between $.02$ and $.025$ we observe a reduction in the number of states in the abstract \ac{MDP} while value is fully maintained. After $.025$, increased reduction in state space size comes at a cost of value. Lastly, as $\epsilon$ is increased in the Random domain, there is a smooth reduction in the number of abstract states with a corresponding cost in the value of the derived policy.

Our experimental results also highlight a noteworthy characteristic of approximate state abstraction in goal-based \acp{MDP}. Taxi exhibits relative stability in state space size and behavior for epsilon up to $.02$ until both fall off dramatically. We attribute the sudden fall off of these quantities to its goal-based nature; once information critical for achieving optimal behavior is lost in the state aggregation, solving the goal -- and so acquiring any reward -- is unattainable. Conversely, in the Random domain, a great deal of near-optimal policies are available to the agent. Thus, even the information for optimal behavior is lost with an increase in $\epsilon$ there are a great deal of near-optimal policies available to the agent which are not lost by the abstraction.

% Figure: Epsilon vs. #States and Epsilon vs Abstract Pol. For NChain and Minefield.
\begin{figure}
\label{fig:main_empirical_results1}
\centering
